\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-04-20 2-the-cubble-pkg.Rmd
a new skeleton of intro incorporating various past writing
why not nest_by

# The cubble package

@@

<!-- ## Why not just use dplyrnest_by -->

<!-- ```{r eval = FALSE} -->
<!-- library(cubble) -->
<!-- library(tidyverse) -->
<!-- left_join(cubble::stations, cubble::climate) |>  -->
<!--   dplyr::nest_by(id, lat,  long,  elev, name, wmo_id) |>  -->
<!--   tidyr::unnest() |>  -->
<!--   group_by(........, geometry) -->
<!-- ``` -->

<!-- ```{r eval = FALSE} -->
<!-- as_cubble(list(spatial = stations, temporal = climate), -->
<!--           key = id, index = date, coords = c(long, lat)) -->
<!-- ``` -->

<!-- ```{r eval = FALSE} -->
<!-- usmap::us_map("states") -->
<!-- ``` -->

>>>>

# The cubble package {#cubble}

@@

<!-- https://www.jstatsoft.org/style#my-latex-paper-does-not-compile-when-there-is-jss-markup-in-section-titles-what-should-i-do -->
\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}}

It may naturally raise the question that why bother a new data structure since the nested form can already be created by the \pkg{dplyr} function \code{nest_by()}?  While \code{nest_by()} and \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data. While `nest_by()` solves this problem on the spatial side, it doesn't solve the temporal side. For temporal data, a combined representation with spatial variables is useful for spatio-temporal operations but its duplication can be a burden to operate on. Also, spatial variables are dropped when summarised and you will need to either specified all of them in the \code{group_by()} before \code{summarise}, or add a joining after \code{summarise()}.

<!-- ```{r} -->
<!-- library(dplyr) -->
<!-- vars <- c(id, lat, long, elev, name, wmo_id) -->
<!-- left_join(cubble::stations, cubble::climate) |>  -->
<!--   group_by(id, lat, long, elev, name, wmo_id) |>  -->
<!--   summarise(tmax_mean = mean(tmax)) -->

<!-- cubble::climate |>  -->
<!--   mutate(month = lubridate::month(date)) |>  -->
<!--   group_by(id, month) |>  -->
<!--   summarise(tmax_mean = mean(tmax)) |>  -->
<!--   ungroup() |>  -->
<!--   left_join(cubble::stations) |>  -->
<!--   nest_by(id, lat, long, elev, name, wmo_id) -->

<!-- ``` --

<!-- ``` r -->
<!-- library(cubble) -->
<!-- library(tidyverse) -->
<!-- left_join(cubble::stations, cubble::climate) |> -->
<!--   dplyr::nest_by(id, lat,  long,  elev, name, wmo_id) |> -->
<!--   tidyr::unnest() |> -->
<!--   group_by(........, geometry) -->
<!-- ``` -->

### 2022-04-20 jss/cubble-jss.tex
why not nest_by

\hypertarget{why-not-just-use-dplyrnestby}{%
\subsection{Why not just use dplyrnestby}\label{why-not-just-use-dplyrnestby}}

>>>>

\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}}

It may naturally raise the question that why bother a new data structure since the nested form can already be created by the \pkg{dplyr} function \code{nest_by()}? While \code{nest_by()} and \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data. While \texttt{nest\_by()} solves this problem on the spatial side, it doesn't solve the temporal side. For temporal data, a combined representation with spatial variables is useful for spatio-temporal operations but its duplication can be a burden to operate on. Also, spatial variables are dropped when summarised and you will need to either specified all of them in the \code{group_by()} before \code{summarise}, or add a joining after \code{summarise()}.
